---
title: "ANOVA Homework Helper"
author: "Nicholas Jacob"
date: "12/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
```

# ANOVA

ANOVA is...


1. An experiment is conducted to determine whether there is a differnce among the mean increases in growth produced by five strains (A, B, C, D and E) of growth hormones for plants. The experimental material consists of 20 cuttings of a shrub (all of equal weight), with four cuttings randomly assigned to each of the five different strains. The increases in weight for each cutting are given in the table below.


|         | A  | B  | C  | D  | E  |
|---------|----|----|----|----|----|
| Plant 1 | 19 | 28 | 25 | 13 | 6  |
| Plant 2 | 8  | 27 | 22 | 12 | 7  |
| Plant 3 | 14 | 20 | 24 | 13 | 15 |
| Plant 4 | 11 | 17 | 22 | 16 | 8  |

Let's add this into R.
```{r cars}
data = data.frame("A" = c(19,8,14,11), "B" = c(28,27,20,17), "C" = c(25,22,24,22), "D" = c(13,12,13,16),"E" = c(6,7,15,8), row.names = 1:4)
data
```


```{r}
aov(1~A,data)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
