---
title: "Non-Parametric Homework"
author: "Nicholas Jacob"
date: "11/24/2020"
output: html_document
---

```{r setup, include = F}
library(boot)
library(caret)
set.seed(42)
```
You may use this document as a starting point for this assignment.  Your submission should be written in R and include all points addressed below.

# Wilcoxson Ranked Sign Test

To keep this assignment simple, we are going to use the built in dataset **ChickWeight** that is included with R.  
```{r ChickWeight}
head(ChickWeight)
```
1. Chickens that do not put on weight early in life are most likely to die.  Examine the hypothesis that diet 1 will not have early weight gain of the chicks. Use a non-parametric method for matched pairs.

<!-- I have done some of the necessary data wrangling for you. -->
<!-- ```{r} -->
<!-- df <- ChickWeight[which(ChickWeight$Time %in% c(0,2) & ChickWeight$Diet %in% c(1)),] -->
<!-- df <- reshape(df, idvar = c("Chick", "Diet"), timevar = "Time", direction = "wide") -->
<!-- df["Difference"] = df$weight.2 - df$weight.0 -->
<!-- df -->
<!-- ``` -->
<!-- The dataframe **df** contains all the necessary info in **Difference**.  You just need to run the test. -->


# Cross Validation

The multiple linear regression using carat, depth, and table to predict price is done below.  

```{r }
fit <- lm(price ~ carat +depth + table, data = diamonds)
summary(fit)
```
2. Repeat this linear model using 10 fold cross validation.  Explain in words what you are doing.  Examine one of the folds carefully explaining the steps involved.  Examine the $R^2$ value and residual mean standard error.  Compare the values you get to the original model.

