---
title: "Applied_Stats_Project"
author: "Nicholas Jacob"
date: "10/22/2020"
output: html_document
---

```{r setup}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
data = read.csv('https://docs.google.com/spreadsheets/d/e/2PACX-1vSDpJqmVSks0f4vLzzcmcTfPJ8TSu4ziCNpTFy_fIY6LibZksRXzCfJYXj9qZd4NiofejxoYSkmLMwu/pub?output=csv')

```

## Olympic Medals

I discovered this data set at <https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results>.  In order to create my data file, I clicked the icon to upload to Google sheets, published it to the web as a csv and copied a link.  Some data will be possible to upload to github but if the file is large github will not do it.  You need the data to be in a raw form, with csv.


```{r}
summary(data)
```
I wonder what sports have had the most athletes of all time.
```{r athletes}
ggplot(data, aes(Sport)) +
  geom_bar()
```
```{r}
table(data$Sport)
```
The table gave me a better representation.  Clearly I could not tell in the bar chart what 
## Scatter Plot

Let's compare height and weight as we will surely find some correlation there!

```{r scatter, echo=FALSE}
ggplot(data, aes(Height, Weight, colour = Sex)) + 
  geom_point() +
  geom_smooth(method="lm")
```

I stuck the color in here to see the differences between the sexes.  Finding that there most likely will be a significant difference in the regression between the sexes, I added the regression lines as well, 'lm' is the linear method.
