---
title: "Applied_Stats_Project"
author: "Nicholas Jacob"
date: "10/22/2020"
output: html_document
---

```{r setup}
library(ggplot2)
library(ggExtra)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
data = read.csv('https://docs.google.com/spreadsheets/d/e/2PACX-1vSDpJqmVSks0f4vLzzcmcTfPJ8TSu4ziCNpTFy_fIY6LibZksRXzCfJYXj9qZd4NiofejxoYSkmLMwu/pub?output=csv')

```

## Olympic Medals

I discovered this data set at <https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results>.  In order to create my data file, I clicked the icon to upload to Google sheets, published it to the web as a csv and copied a link.  Some data will be possible to upload to github but if the file is large github will not do it.  You need the data to be in a raw form, with csv.


```{r}
summary(data)
```
I wonder what sports have had the most athletes of all time.
```{r athletes}
ggplot(data, aes(Sport)) +
  geom_bar()
```
```{r}
table(data$Sport)
```
The table gave me a better representation.  Clearly I could not tell in the bar chart what the games were.  Let me try one more way and include relative frequency.

```{r}
#options(digits=2)
freq <- rbind(table(data$Sport),table(data$Sport)/length(data$Sport))
row.names(freq)<-c('Frequency','Relative Frequency')
freq
```
meh, that is still not great.  One more time?
```{r}
percent <- function(x, digits = 2, format = "f", ...) {      # Create user-defined function
  paste0(formatC(x * 100, format = format, digits = digits, ...), "%")
} 
data %>%
  group_by(Sport) %>%
  summarize(Frequency = length(Sport),RelativeFrequency = percent(length(Sport)/271116))
```

Such a nicer, usable table!  I needed the tidyverse library to do this but I think it was worth it.  You should also notice that I created a function for doing the percent too!  Take a minute and try to understand what that function is doing because somethings will need to be modified to make it work for your data!

Let's look at a couple of the sports.  Athletics, cycling and rowing are of interest to me.

```{r}
sports <- c('Athletics','Cycling', 'Rowing')

ggplot(subset(data, Sport %in% sports), aes(Sport,fill = Sex)) +
  geom_bar()
```
Wow so many more men than women!  Let's see if the number of medals is as far apart.
```{r}
table(data$Medal,data$Sex)
```
Clearly there have been way more men's medals awarded then women's medals.  Let's see if we can figure out what year women start competing at the same level as the men.
```{r}
table(data$Sex,data$Year)
```
Interestingly the female contingent is still smaller than the male.  I wonder if there are less medals awarded too.
```{r}
year16 <- data[which(data$Year == 2016),]
table(year16$Sex,year16$Medal)
```
Yes, less medals all the way around in the most recent year.  

I guess I should do nationality counts and look at the number of medals won too!
```{r}
data %>%
  group_by(Medal,Team,Sex) %>%
  summarize(Medals = length(Medal))
```
Way too much data there to synthesize!

Let's look at age and Team.
```{r}
data %>%
  group_by(Team) %>%
  summarize(AveargeAge = mean(Age, na.rm = TRUE))
```
Okay well again there is just too many teams!  I wonder if there is a gender difference in age?
```{r}
data %>%
  group_by(Sex) %>%
  summarize(AverageAge = mean(Age, na.rm = TRUE), StandardDeviation = sd(Age, na.rm = TRUE), SampleSize = length(Age))
```
I bet those are statistically significant differences! 

```{r AgeTest}
t.test(data$Age~data$Sex)
```
We see that yes indeed the differences in ages are significant!  Lets visualize this too!
```{r BoxSexAge}
ggplot(data = data,aes(x= Age, y =Sex))+
  geom_boxplot()
```

Seriously who are all these ancient olympians?

```{r}
outliers <- boxplot.stats(data$Age)$out #finds the outliers
oldies <- data[which(data$Age %in% c(outliers)),] #grabs all the data from the outliers and renames it

oldies %>%
  group_by(Sport) %>%
  summarize(AverageAge = mean(Age), MaxAge = max(Age), MinAge = min(Age))
```
Well the 97 year old participated in 'Art Competitions'...  This table is misleading in that I am only looking at the outliers!  There is also a 10 year old gymnast!

## Scatter Plot

Let's compare height and weight as we will surely find some correlation there!

```{r scatter, echo=FALSE}
g <- ggplot(data, aes(Height, Weight, colour = Sex)) + 
  geom_point() +
  geom_smooth(method="lm")
ggMarginal(g, type = "histogram")
```

I stuck the color in here to see the differences between the sexes.  Finding that there most likely will be a significant difference in the regression between the sexes, I added the regression lines as well, 'lm' is the linear method.  The histograms on the outside are all about looking at the normality of the data. 

I will check the normality in another way.  QQ-Plots!

```{r}
ggplot(data, aes(sample = Height)) + 
  stat_qq() +
  stat_qq_line()
```
```{r}
ggplot(data, aes(sample = Weight)) + 
  stat_qq() +
  stat_qq_line()
```
QQ-Plots are all about looking at the normality of the datasets.  Height fits very well on the line but Weight does not!  I wonder if it will do better by restricting the genders.
```{r}
ggplot(data[ which(data$Sex == 'F'),],aes(sample = Weight)) +
  stat_qq() +
  stat_qq_line()
```

```{r}
ggplot(data[ which(data$Sex == 'M'),],aes(sample = Weight)) +
  stat_qq() +
  stat_qq_line()
```

Both deviate from the line.  Essentially that suggests that our data is not normal.