---
title: "ScatterPlots"
author: "Nicholas Jacob"
date: "11/13/2020"
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(ggExtra)
library(tidyverse)
library(GGally)

data = read.csv('https://docs.google.com/spreadsheets/d/e/2PACX-1vSDpJqmVSks0f4vLzzcmcTfPJ8TSu4ziCNpTFy_fIY6LibZksRXzCfJYXj9qZd4NiofejxoYSkmLMwu/pub?output=csv')

```

## Scatter Plot

Let's compare height and weight as we will surely find some correlation there!

```{r scatter, echo=FALSE}
g <- ggplot(data, aes(Height, Weight, colour = Sex)) + 
  geom_point() +
  geom_smooth(method="lm")
ggMarginal(g, type = "histogram")
```

I stuck the color in here to see the differences between the sexes.  Finding that there most likely will be a significant difference in the regression between the sexes, I added the regression lines as well, 'lm' is the linear method.  The histograms on the outside are all about looking at the normality of the data. 

I will check the normality in another way.  QQ-Plots!

```{r}
ggplot(data, aes(sample = Height)) + 
  stat_qq() +
  stat_qq_line()
```
```{r}
ggplot(data, aes(sample = Weight)) + 
  stat_qq() +
  stat_qq_line()
```
QQ-Plots are all about looking at the normality of the datasets.  Height fits very well on the line but Weight does not!  I wonder if it will do better by restricting the genders.
```{r}
ggplot(data[ which(data$Sex == 'F'),],aes(sample = Weight)) +
  stat_qq() +
  stat_qq_line()
```

```{r}
ggplot(data[ which(data$Sex == 'M'),],aes(sample = Weight)) +
  stat_qq() +
  stat_qq_line()
```

Both deviate from the line.  Essentially that suggests that our data is not normal.

I am going to continue anyway but I should be wary of my statistics, in particularly my $p$ value!

```{r}
fit <- lm(Weight ~ Height + Sex, data)
summary(fit)
```
All the variables are significant!  Being male adds about 5 kilos!


I don't know about you but I am interested to see if the year makes a difference.  Maybe we are breeding larger athletes?  I'll tack on sport too!
```{r multipleRegression}
fit <- lm(Weight ~ Height + Sex + Year + Sport, data)
summary(fit)
```

Year is not statistically significant.  Interestingly some sports make a big difference and a few don't matter!  Can you predict which ones will matter and why?
