---
title: "Bootstrap and Cross Validation Homework"
author: "Nicholas Jacob"
date: "11/24/2020"
output: html_document
---

```{r setup, include = F}
library(boot)
library(caret)
set.seed(42)
```
You may use this document as a starting point for this assignment.  Your submission should be written in R and include all points addressed below.

# Bootstrap

To keep this assignment simple, we are going to use the built in dataset **diamonds** that is included with R.  
```{r diamonds}
summary(diamonds)
```
1. A quick google search suggested that I should purchase a 1 carat diamond as an engagement ring.  Using bootstrapping, test the hypothesis that the mean is 1 carat.  Explain in words how to create a bootstrap and how to create a bootstrap distribution for the mean.  Make sure to state the hypothesis, express a confidence interval, $p$ value, and state the conclusion in the proper statistical terms for the mean.



# Cross Validation

The multiple linear regression using carat, depth, and table to predict price is done below.  

```{r }
fit <- lm(price ~ carat +depth + table, data = diamonds)
summary(fit)
```
2. Repeat this linear model using 10 fold cross validation.  Explain in words what you are doing.  Examine one of the folds carefully explaining the steps involved.  Examine the $R^2$ value and residual mean standard error.  Compare the values you get to the original model.

