---
title: "Non-Parametric Homework"
author: "Nicholas Jacob"
date: "11/24/2020"
output: html_document
---

```{r setup, include = F}

```
You may use this document as a starting point for this assignment.  Your submission should be written in R and include all points addressed below.


To keep this assignment simple, we are going to use the built in dataset **ChickWeight** that is included with R.  
```{r ChickWeight}
head(ChickWeight)
```


# Wilcoxson Ranked Sign Test


1. Chickens that do not put on weight early in life are most likely to die.  Examine the hypothesis that diet 1 will not have early weight gain of the chicks. Use a non-parametric method for matched pairs.

<!-- The code below does most of the necessary data wrangling for you.  I've commented this out because I wanted you to try this step but if you cannot figure it out it is here... -->
<!-- ```{r} -->
<!-- df <- ChickWeight[which(ChickWeight$Time %in% c(0,2) & ChickWeight$Diet %in% c(1,2)),] -->
<!-- df <- reshape(df, idvar = c("Chick", "Diet"), timevar = "Time", direction = "wide") -->
<!-- df["Difference"] = df$weight.2 - df$weight.0 -->
<!-- df -->
<!-- ``` -->
<!-- The dataframe **df** contains all the necessary info in **Difference**.  You just need to run the test. -->


# Kruskal-Wallis

We are going to repeat the test above but examine if there is a difference in the **diets**.


2. Examine the claim that there is a difference in early weight gain between the different diets.  Use a non-parametetric method for comparing multiple medians.  Be sure to include an appropriate visualization.

<!-- The code below does the data wrangling for you, but you should try it before uncommenting! -->
<!-- ```{r DrJacobsWrangle} -->
<!-- df2 <- ChickWeight[which(ChickWeight$Time %in% c(0,2)),] -->
<!-- df2 <- reshape(df2, idvar =c("Chick","Diet"),timevar = "Time",direction = "wide") -->
<!-- df2["Difference"] = df2$weight.2 - df2$weight.0 -->
<!-- df2 -->
<!-- ``` -->

