---
title: "ANOVA Project"
author: "Nicholas Jacob"
date: "12/10/2020"
output: html_document
---


```{r setup, include=FALSE}
library(ggplot2)
data = read.csv('https://docs.google.com/spreadsheets/d/e/2PACX-1vSDpJqmVSks0f4vLzzcmcTfPJ8TSu4ziCNpTFy_fIY6LibZksRXzCfJYXj9qZd4NiofejxoYSkmLMwu/pub?output=csv')
```

## Olympics

Keeping with my project, I am now going to explore the olympic dataset and preform an few ANOVA computations.  Let's ask if the sport they play affects their height.  Really I am asking is the mean height of all athletes across sports the same.

$$
H_0: \mu_1 = \mu_2=\cdots=\mu_n\\
H_A: \mu_i\neq\mu_j \text{ for some }i\text{ and }j
$$

```{r}
a <- aov(Height ~ Sport , data = data)
summary(a)
```
Clearly there is some difference, so we will reject the null hypothesis.  Let's explore the assumptions of the test.
```{r}
plot(a,1)
```
The red line sits right in the middle here so eye balling it, it looks like the assumption of equal variances is okay.
```{r}
plot(a,2)
```

The QQ-Plot also looks acceptable suggesting the data is normal.  So I feel confident rejecting the null hypothesis and saying that there is evidence that the mean height of athletes in different sports do have a different.

```{r}
y1 <-  mean(data$Height, na.rm = TRUE)
ggplot(data = data, aes(x = Sport, y = Height))+
  geom_jitter(color = 'grey') +
  stat_summary(fun.data = 'mean_se', color = "red") +
  geom_hline(yintercept = y1,  color = "blue",linetype = "dashed")
```

# Two Way ANOVA

Next I wonder if gender plays a significant role here too?  I'll add it to my ANOVA and explore.  There will be three hypotheses here.
```{r}
a2 <- aov(Weight ~ Sport*Sex, data)
summary(a2)
```


