---
title: "Project Proposal"
author: "Nicholas Jacob"
date: "1/19/2021"
output: html_document
---


```{r setup}
data = read.csv('https://docs.google.com/spreadsheets/d/e/2PACX-1vSDpJqmVSks0f4vLzzcmcTfPJ8TSu4ziCNpTFy_fIY6LibZksRXzCfJYXj9qZd4NiofejxoYSkmLMwu/pub?output=csv')

```

# Olympic Medals Data Exploration

For this first assignment, I want you to get into R and load your data!


I discovered this data set at <https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results>.  In order to create my data file, I clicked the icon to upload to Google sheets, published it to the web as a csv and copied a link.  Some data will be possible to upload to github but if the file is large github will not do it.  You need the data to be in a raw form, with csv.

Once you have your data loaded up into R, lets do two things with it!  First I am going to show a snippet of the data with the `head` command.  Create an R environment using three tick marks (by the tilde) and an `{r}` and close it out with three more tick marks.  Anything inside can be run as R code (anything outside is just text explaining your project) 

```{r}
head(data)
```

So we see a few athletes from different times and events.  Notice that Aaftink is repeated, might worry about that later.  Lastly I'll run some real quick statistics.


```{r}
summary(data)
```

This went through all the columns and did summary statistics on the quantitative variables.  Okay that is all I want for the first assignment!

All other parts of the project can be found on the following series of links.

0. [This Document](https://rpubs.com/nurfnick/715221)
1. [Exploratory Data Analysis](https://rpubs.com/nurfnick/751605)
2. [Hypothesis Testing](https://rpubs.com/nurfnick/715229)
3. [Regression](https://rpubs.com/nurfnick/720617)
4. [Bootstrapping](https://rpubs.com/nurfnick/696714)
5. [Cross Validation](https://rpubs.com/nurfnick/751613)
6. [Contingency Tables and Goodness of Fit](https://rpubs.com/nurfnick/730813)
7. [ANOVA](https://rpubs.com/nurfnick/733672)
8. [NonParametric Methods](https://rpubs.com/nurfnick/743956)